---
title: "Analysis code for 'Life history traits, pace-of-life and dispersal among and within five species of *Trichogramma*: a comparative analysis'"
author: "Chloé GUICHARNAUD, Géraldine Groussier, Erwan Beranger, Laurent Lamy, Maxime Dahirel, Elodie Vercken"
date: 
output:
  html_document:
  df_print: paged
  html_notebook: default
  pdf_document: default
  editor_options:
    chunk_output_type: inline
  chunk_output_type: inline
---

## Download necessary libraries

```{r load-packages, warning= FALSE, message = FALSE}
library(here)      # CRAN 1.0.1

library(tidyverse) # CRAN 1.3.2

library(cmdstanr)  # [github::stan-dev/cmdstanr] 0.5.2 ## Stan backend
library(brms)      # CRAN 2.17.0
library(bayesplot) # CRAN 1.9.0
library(tidybayes) # CRAN 3.0.2
library(patchwork) # CRAN 1.1.2
library(ggtext)    # CRAN 0.1.2 

options(mc.cores = 4)

here::i_am("R/script_Trichogramma_POL_rmd.Rmd")
```

## Data loading and prep

### Import dataset

```{r load-data}

data_Trichogramma <- read_csv(here("data","data_Trichogramma.csv"))
```

This dataset contains:

-   `species` : name of the *Trichogramma* species (one of 5)
-   `line` : name of the *Trichogramma* line (one of 32)
-   `replicate` : replicate number among lines
-   `type` : `F` for single-vial system, `D` for two-vial systems
-   `block` : temporal blocks
-   `date_start` : date when females where placed in the experimental system
-   `date_emerging` : first emerging descendant observed date
-   `hour_emerging` : first emerging descendant observed hour
-   `total_hours` : observed time in hour between earliest start of the experiment (date_start, at 9 am) and observed emerging hour
-   `lowbound` : low bound of observed developmental time interval (in days)
-   `upbound` : up bound of observed developmental time interval (in days)
-   `dispersal_status` : `null_disp` for single-vial system, no dispersal possible `yes_disp` for dispersed replicates in two-vial systems `no_disp` for non-dispersed replicates in two-vial systems
-   `nb_egg` : number of parazitized host eggs counted

### Adding columns for running models

To run our models, we need to add some information to our dataset:

-   For development time observations, precise the type of censoring (here, interval censoring for everything)
-   Transform `dispersal_status` into something that says in model-understandable terms if a valid dispersal observation was made and the result (0,1,NA)
-   Because the 2-vial data are not used to analyse fecundity and development time, add variables that will be used in `brm()` along with `subset()` to specify which row is valid to use in the model for each variable of interest

```{r data-update}

# Add censoring info
data_Trichogramma$censore <- "interval"

# Clarify dispersal
data_Trichogramma <- data_Trichogramma |>
  mutate(dispersal = case_when(
    dispersal_status == "yes_disp" ~ 1,
    dispersal_status == "no_disp" ~ 0,
    dispersal_status == "null_disp" ~ NA_real_,
    TRUE ~ NA_real_
  ))

# Add valid subsetting variables for development, dispersal and fecundity.

data_Trichogramma <- data_Trichogramma |>
  mutate(
    # development times are usable here if they come from the 1 vial experiments and have non-NA emergence dates
    valid_development = case_when(
      is.na(date_emerging) == FALSE & type == "F" ~ 1,
      TRUE ~ 0
    ),
    # fecundities are usable if they come from the 1 vial experiments and have non-NA egg numbers
    valid_fecundity = case_when(
      is.na(nb_egg) == FALSE & type == "F" ~ 1,
      TRUE ~ 0
    ),
    # dispersal are only valid in the 2-vial systems
    valid_dispersal = case_when(
      is.na(dispersal) == FALSE ~ 1,
      TRUE ~ 0
    ),
  )
```

## Models

### Choice of families

Dispersal data are binary, so a Bernoulli model is really the only choice. Development times are time to event data, so a lognormal model should be appropriate at least at first. Fecundity are counts, so a Poisson model should be a good start. However:

```{r}
hist(subset(data_Trichogramma$nb_egg, data_Trichogramma$valid_fecundity == 1))
mean(subset(data_Trichogramma$nb_egg, data_Trichogramma$valid_fecundity == 1))
var(subset(data_Trichogramma$nb_egg, data_Trichogramma$valid_fecundity == 1))
```

there is some hints of overdispersion and/or zero-inflation. These are not conclusive (it could be high between-line heterogeneity) but worth considering.

### prior setting

To run our models, we need to specify priors for our three response variables. Priors for random effects and correlation matrices will be standard ones. Priors for the fixed-effects intercepts (species-specific or not) will also be standard for the dispersal models (since these are binary data on a standard 0-1 scale); but they need to be scaled for the fecundity and development time data. We suggest to use the median values as a scaling factor for the priors:

```{r median-values}

# development time
median(subset(data_Trichogramma$lowbound, data_Trichogramma$valid_development == 1)) # = 11

# fecundity
median(subset(data_Trichogramma$nb_egg, data_Trichogramma$valid_fecundity == 1)) # = 19
```

### Family choice for fecundity

We tried a first univariate negative binomial model for fecundity:

```{r fecundity-inflated, warning= FALSE, message= FALSE, results = 'hide'}

bf_fecundity_negbino <- bf(nb_egg | subset(valid_fecundity) ~ 0 + species + (1 | p | line), family = "negbinomial")

prior_fecundity_negbino <- c(
  set_prior("normal(0,1)", class = "sd"),
  set_prior("normal(log(19),1)", class = "b") # the median info is used on the log scale because of the log link for the negbinomial model
)

mod_fecundity_negbino <- brm(
  bf_fecundity_negbino,
  data = data_Trichogramma,
  prior = prior_fecundity_negbino,
  seed = 404,
  warmup = 2000, iter = 4500, chains = 4,
  backend = "cmdstanr"
)
```

There may be some divergences, but we'll let them pass for now, because:

```{r fecundity check}

pp_check(mod_fecundity_negbino, ndraws = 150) # ok for range, but not for peak height
pp_check(mod_fecundity_negbino, "ecdf_overlay", ndraws = 150) # ok for range, but not for peak height
```

Posterior does not correctly fit observations, where two peaks are present, one representing an overabundance of 0. It is especially bad catching the correct number of zeroes. Therefore, a zero-inflated negative binomial function is used in multivariate models.

### Main models

**Model 1**

The first model take into account inter and intra specific line covariation:

```{r}
# priors

prior1 <- c(
  # development prior
  set_prior("normal(log(11),1)", class = "b", resp = "lowbound"),
  set_prior("normal(0,1)", class = "sd", resp = "lowbound"),
  set_prior("normal(0,1)", class = "sigma", resp = "lowbound"),

  # fecundity prior
  set_prior("normal(log(19),1)", class = "b", resp = "nbegg"),
  set_prior("normal(0,1.5)", class = "b", dpar = "zi", resp = "nbegg"),
  set_prior("normal(0,1)", class = "sd", resp = "nbegg"),
  set_prior("normal(0,1)", class = "b", nlpar = "invshape", lb = 0, resp = "nbegg"),

  # dispersal prior
  set_prior("normal(0,1.5)", class = "b", resp = "dispersal"),
  set_prior("normal(0,1)", class = "sd", resp = "dispersal"),

  # covariation
  set_prior("lkj(2)", class = "cor")
)
```

```{r}
# formulas

bf_fecundity_0inflated1 <- bf(nb_egg | subset(valid_fecundity) ~ 0 + species + (1 | p | line), # fecundity without 0 inflation
  zi ~ 0 + species + (1 | p | line), # zi: new variable for nbegg=0
  nlf(shape ~ 1 / invshape), # nlf = specify non linear formula to brms
  invshape ~ 1,
  # use invshape so generic priors are easier to set (see https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations, the case of negative binomial)
  family = "zero_inflated_negbinomial"(link_shape = "identity")
) # fixing the link for shape to identity despite the use of a formula (which would otherwise make it default to log link) is important for correct prior setting

bf_dvlpmt_cens1 <- bf(lowbound | cens(censore, upbound) + subset(valid_development) ~ 0 + species + (1 | p | line), family = "lognormal")

bf_dispersal1 <- bf(dispersal | subset(valid_dispersal) ~ 0 + species + (1 | p | line), family = "bernoulli")
```

```{r multivariate model 1, warning= FALSE, message= FALSE, results = 'hide'}

if (file.exists(here("R_output", "modeld1.Rdata"))) {
  load(here("R_output", "model1.Rdata"))
} else {
  model1 <- brm(
    mvbf(bf_fecundity_0inflated1 + bf_dvlpmt_cens1 + bf_dispersal1),
    data = data_Trichogramma,
    prior = prior1,
    seed = 404,
    warmup = 2000, iter = 4500, chains = 4,
    backend = "cmdstanr"
  )

  save(list = "model1", file = here("R_output", "model1.Rdata"))
}
```

**Model 2**

To be able to more fully describe covariation patterns between lines, the second model removes species-level effects, so removes the partition that was made between within-species and among-species line level covariation

```{r}
# priors

prior2 <- c(
  # development prior
  set_prior("normal(log(11),1)", class = "Intercept", resp = "lowbound"),
  set_prior("normal(0,1)", class = "sd", resp = "lowbound"),
  set_prior("normal(0,1)", class = "sigma", resp = "lowbound"),

  # fecundity prior
  set_prior("normal(log(19),1)", class = "Intercept", resp = "nbegg"),
  set_prior("normal(0,1.5)", class = "Intercept", dpar = "zi", resp = "nbegg"),
  set_prior("normal(0,1)", class = "sd", resp = "nbegg"),
  set_prior("normal(0,1)", class = "b", nlpar = "invshape", lb = 0, resp = "nbegg"),


  # dispersal prior
  set_prior("normal(0,1.5)", class = "Intercept", resp = "dispersal"),
  set_prior("normal(0,1)", class = "sd", resp = "dispersal"),

  # covariation
  set_prior("lkj(2)", class = "cor")
)
```

```{r}
# formula

bf_fecundity_0inflated2 <- bf(nb_egg | subset(valid_fecundity) ~ (1 | p | line),
  zi ~ (1 | p | line),
  nlf(shape ~ 1 / invshape),
  invshape ~ 1,
  family = "zero_inflated_negbinomial"(link_shape = "identity")
)

bf_dvlpmt_cens2 <- bf(lowbound | cens(censore, upbound) + subset(valid_development) ~ (1 | p | line), family = "lognormal")

bf_dispersal2 <- bf(dispersal | subset(valid_dispersal) ~ (1 | p | line), family = "bernoulli")
```

```{r multivariate model 2, warning= FALSE, message= FALSE, results = 'hide'}


if (file.exists(here("R_output", "model2.Rdata"))) {
  load(here("R_output", "model2.Rdata"))
} else {
  model2 <- brm(
    mvbf(bf_fecundity_0inflated2 + bf_dvlpmt_cens2 + bf_dispersal2),
    data = data_Trichogramma,
    prior = prior2,
    seed = 404,
    warmup = 2000, iter = 4500, chains = 4,
    backend = "cmdstanr"
  )

  save(list = "model2", file = here("R_output", "model2.Rdata"))
}
```

### various model checks and diagnostics

Now, let's quickly check models' results

```{r}
summary(model1) # check for ESS and Rhat
summary(model2)
```

```{r}
# convergence/stationarity checks
plot(model1) # MCMC seem to overlap nicely
plot(model2)
```

```{r post model check, warning= FALSE, message= FALSE }

# fecundity:
pp_check(model1, resp = "nbegg", ndraws = 150) # follows the curve but fails to correctly present the dip between ZI and mean fecundity
# but it does capture the correct % of zeroes
## note: if you re-run the model using a ZI Poisson instead, the dip is better but the model performs way worse in other aspects (model comparison using WAIC/LOO), reflecting existing overdispersion. So that's probably as good as one can get without more information (prior or data)
pp_check(model2, resp = "nbegg", ndraws = 150)
pp_check(model1, resp = "nbegg", type = "ecdf_overlay", ndraws = 150)
pp_check(model2, resp = "nbegg", type = "ecdf_overlay", ndraws = 150)
pp_check(model1, resp = "nbegg", type = "stat_2d")
pp_check(model2, resp = "nbegg", type = "stat_2d")


# dispersal (but due to binary type, unlikely to be very informative, most checks should always perform well)
pp_check(model1, resp = "dispersal", ndraws = 150)
pp_check(model2, resp = "dispersal", ndraws = 150)

# development time

# there are no good simple way to use pp_check for interval censored data.
# but we can get *some* information by being creative

source <- data_Trichogramma |>
  filter(valid_development == 1)

yrep1 <- posterior_predict(model1, resp = "lowbound", summary = FALSE) # posterior predictive distribution of data

ppc_intervals_data(source$lowbound, yrep = yrep1) |>
  mutate(lowbound_obs = y_obs, upbound_obs = source$upbound) |>
  arrange(lowbound_obs) |>
  mutate(x = row_number()) |>
  ggplot() +
  geom_ribbon(aes(x, ymin = lowbound_obs, ymax = upbound_obs), fill = "blue", alpha = 0.5) +
  geom_pointinterval(aes(x, y = m, ymin = ll, ymax = hh)) +
  scale_y_continuous("development time") +
  scale_x_continuous("rank order (observed)")

# in this plot the ribbon denotes the observed data (the 2 day interval)
# and the points and error bars represents the posterior mean prediction and CI95 (quantile interval)
# we can see that the model does a decent job at matching the data (not many points are outside the observed range, and for the majority of the predictions where the point/mean is outside the observed range, the CI is comfortably overlapping it)

yrep2 <- posterior_predict(model2, resp = "lowbound", summary = FALSE) # posterior predictive distribution of data

ppc_intervals_data(source$lowbound, yrep = yrep2) |>
  mutate(lowbound_obs = y_obs, upbound_obs = source$upbound) |>
  arrange(lowbound_obs) |>
  mutate(x = row_number()) |>
  ggplot() +
  geom_ribbon(aes(x, ymin = lowbound_obs, ymax = upbound_obs), fill = "blue", alpha = 0.5) +
  geom_pointinterval(aes(x, y = m, ymin = ll, ymax = hh)) +
  scale_y_continuous("development time") +
  scale_x_continuous("rank order (observed)")

# Building on this, one can probably think of other ways to check that the model does well for development time, despite the interval censoring barrier
```

## predictions and plots

### generation

We create posterior datasets for the 2 models

```{r post-model dataset, warning= FALSE, message= FALSE}
# Create posterior dataset for the first model

unique_lines <- data_Trichogramma |>
  select(species, line) |>
  distinct() |>
  mutate(valid_development = 1, valid_dispersal = 1, valid_fecundity = 1) # need all subset columns

modcens1_data_dvlp <- unique_lines |>
  add_epred_draws(model1, resp = "lowbound") |>
  rename(development_time = .epred)


modcens1_data_disp <- unique_lines |>
  add_epred_draws(model1, resp = "dispersal") |>
  rename(dispersal_rate = .epred)


modcens1_data_fec <- unique_lines |>
  add_epred_draws(model1, resp = "nbegg", dpar = c("zi", "mu")) |>
  rename(fecundity = .epred)


model1_data <- modcens1_data_dvlp |>
  left_join(modcens1_data_disp) |>
  left_join(modcens1_data_fec) |>
  subset(select = -c(.iteration, .chain, .row, valid_development, valid_dispersal, valid_fecundity))


modcens2_data_dvlp <- unique_lines |>
  add_epred_draws(model2, resp = "lowbound") |>
  rename(development_time = .epred)


modcens2_data_disp <- unique_lines |>
  add_epred_draws(model2, resp = "dispersal") |>
  rename(dispersal_rate = .epred)


modcens2_data_fec <- unique_lines |>
  add_epred_draws(model2, resp = "nbegg", dpar = c("zi", "mu")) |>
  rename(fecundity = .epred)


model2_data <- modcens2_data_dvlp |>
  left_join(modcens2_data_disp) |>
  left_join(modcens2_data_fec) |>
  subset(select = -c(.iteration, .chain, .row, valid_development, valid_dispersal, valid_fecundity))
```

Here, only fixed effect posterior are kept to have species-level trait values

```{r posterior fixed effects only}

unique_species <- data_Trichogramma |>
  select(species) |>
  distinct() |>
  mutate(valid_development = 1, valid_dispersal = 1, valid_fecundity = 1) # need all subset columns

# Create posterior dataset for the first model

fixed_data_dvlp <- unique_species |>
  mutate(valid_development = 1, valid_dispersal = 1, valid_fecundity = 1) |> # need all subset columns
  add_epred_draws(model1, resp = "lowbound", re_formula = NA) |>
  rename(development_time = .epred)


fixed_data_disp <- unique_species |>
  mutate(valid_dispersal = 1, valid_fecundity = 1, valid_development = 1) |>
  add_epred_draws(model1, resp = "dispersal", re_formula = NA) |>
  rename(dispersal_rate = .epred)


fixed_data_fec <- unique_species |>
  mutate(valid_fecundity = 1, valid_development = 1, valid_dispersal = 1) |>
  add_epred_draws(model1, resp = "nbegg", dpar = c("zi", "mu"), re_formula = NA) |>
  rename(fecundity = .epred)


fixed_data <- fixed_data_dvlp |>
  left_join(fixed_data_disp) |>
  left_join(fixed_data_fec) |>
  subset(select = -c(.iteration, .chain, .row, valid_development, valid_dispersal, valid_fecundity))
```

### Table 1: species level (fixed effects) predictions, summary table

```{r traits values}
## fixed trait table, fixed effect only (table 1)

table_trait_values_development <- fixed_data |>
  group_by(species) |>
  mean_hdi(development_time) |>
  mutate(development_time = paste(round(development_time, 2), " [", round(.lower, 2), "; ", round(.upper, 2), "]", sep = "")) |>
  select(-c(.lower, .upper, .width, .point, .interval))

table_trait_values_fecundity <- fixed_data |>
  group_by(species) |>
  mean_hdi(mu) |>
  mutate(fecundity_no_retention = paste(round(mu, 2), " [", round(.lower, 2), "; ", round(.upper, 2), "]", sep = "")) |>
  select(-c(.lower, .upper, .width, .point, .interval, mu))

table_trait_values_retention <- fixed_data |>
  group_by(species) |>
  mean_hdi(zi) |>
  mutate(retention_probability = paste(round(zi, 2), " [", round(.lower, 2), "; ", round(.upper, 2), "]", sep = "")) |>
  select(-c(.lower, .upper, .width, .point, .interval, zi))

table_trait_values_dispersal <- fixed_data |>
  group_by(species) |>
  mean_hdi(dispersal_rate) |>
  mutate(dispersal_rate = paste(round(dispersal_rate, 2), " [", round(.lower, 2), "; ", round(.upper, 2), "]", sep = "")) |>
  select(-c(.lower, .upper, .width, .point, .interval))


table1 <- left_join(table_trait_values_development, table_trait_values_fecundity) |>
  left_join(table_trait_values_retention) |>
  left_join(table_trait_values_dispersal)

table1
```

### pairwise differences between species (for "significantly different" markers on table 1)

```{r compare test, eval = FALSE}

species_compare <- fixed_data |>
  pivot_longer(cols = c(development_time, dispersal_rate, fecundity, zi, mu), names_to = "trait") |>
  group_by(trait) |>
  compare_levels(variable = value, by = species) |>
  group_by(trait, species) |>
  mean_hdi()

print(species_compare, n = Inf)

# one can also do it by line (commented out because takes some time, there are a lot of pairwise comparisons between 28 lines)
# lines_compare <- model1_data |>
#  pivot_longer(cols=c(development_time,dispersal_rate,fecundity,zi,mu), names_to = "trait") |>
#  group_by(trait) |>
#  compare_levels(variable=value,by=line) |>
#  group_by(trait,line) |>
#  mean_hdi()
```

### Figure 2: species and line level posterior per trait

```{r plot trait values}
species_palette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#CC79A7")
# colour palette for species

position_lines <- c(0.8,0.9,1.1,1.2,1.8,1.85,1.9,1.95,2.05,2.1,2.15,2.2,2.25,2.85,2.9,2.95,3.05,3.1,3.15,3.2,3.9,4.1,4.2,4.9,4.95,5.05,5.1,5.15)
# plot position for lines' IC

model1_data$line = factor(model1_data$line, levels = c("ISA11967","ISA11969","ISA5544","ISA6646","F3-2","F3-9","F5-11","F5-12","F5-8","F6-4","I2-16","I6-5","PR002","BIO-XA","BIO-XE","BIO-XK","H-03","HY-05","N-05","Q-05","81a","ISA11235","ISA11367","BL110","CRV065","FPV025","FPV034-A","PR007"))


# sorting lines in order of the graphical representation

## plot posterior traits values per species (fixed effects only), with posterior lines values as boxplot
model1_data$full_species <- paste0("*T. ", model1_data$species, "*")
fixed_data$full_species <- paste0("*T. ", fixed_data$species, "*")

## observed means for plotting (approximated for development and fecundity variables)
observed_species_disp <- data_Trichogramma |>
  filter(valid_dispersal == 1) |>
  group_by(species) |>
  summarise(disp = mean(dispersal)) |>
  mutate(full_species = paste0("*T. ", species, "*"))

observed_species_dev <- data_Trichogramma |>
  filter(valid_development == 1) |>
  group_by(species) |>
  summarise(time = mean((upbound + lowbound) / 2)) |>
  mutate(full_species = paste0("*T. ", species, "*"))

observed_species_zi <- data_Trichogramma |>
  filter(valid_fecundity == 1) |>
  group_by(species) |>
  summarise(zi = mean(nb_egg == 0)) |>
  mutate(full_species = paste0("*T. ", species, "*"))

observed_species_mu <- data_Trichogramma |>
  filter(valid_development == 1 & nb_egg > 0) |>
  group_by(species) |>
  summarise(mu = mean(nb_egg)) |>
  mutate(full_species = paste0("*T. ", species, "*"))



plot_dispersal <- ggplot() +
  stat_pointinterval(
    data = model1_data,
    aes(full_species, dispersal_rate, group = line),
    col = "grey70",
    x = position_lines,
    .width = 0.95
  ) +
  stat_eye(
    data = fixed_data,
    aes(full_species, dispersal_rate, fill = species),
    slab_alpha = 0.5,
    .width = c(0.001, 0.95),
    point_interval = "mean_hdi"
  ) +
  geom_point(
    data = observed_species_disp,
    aes(full_species, disp), pch = 21, fill = "white", size = 2
  )  +
  scale_fill_manual(values = species_palette) +
  scale_x_discrete("", guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous("dispersal rate") +
  annotate("text", x=1, y=0.3, label= "a") +
  annotate("text", x=2, y=0.3, label= "a") +
  annotate("text", x=3, y=0.3, label= "a") +
  annotate("text", x=4, y=0.3, label= "a") +
  annotate("text", x=5, y=0.4, label= "a") +
  coord_cartesian(xlim = c(1, 5))


plot_dispersal


plot_development <- ggplot() +
  stat_pointinterval(
    data = model1_data,
    aes(full_species, development_time, group = line),
    col = "grey70",
    x = position_lines,
    .width = 0.95
  ) +
  stat_eye(
    data = fixed_data,
    aes(full_species, development_time, fill = species),
    slab_alpha = 0.5,
    .width = c(0.001, 0.95),
    point_interval = "mean_hdi"
  ) +
  geom_point(
    data = observed_species_dev,
    aes(full_species, time), pch = 21, fill = "white", size = 2
  ) +
  scale_fill_manual(values = species_palette) +
  scale_x_discrete("", guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous("development time (days)") +
  annotate("text", x=1, y=14.9, label= "a") +
  annotate("text", x=2, y=13, label= "b") +
  annotate("text", x=3, y=14, label= "ab") +
  annotate("text", x=4, y=14, label= "ac") +
  annotate("text", x=5, y=13.5, label= "bc") +
  coord_cartesian(xlim = c(1, 5))

plot_development

plot_fecundity <- ggplot() +
  stat_pointinterval(
    data = model1_data,
    aes(full_species, mu, group = line),
    col = "grey70",
    x = position_lines,
    .width = 0.95
  ) +
  stat_eye(
    data = fixed_data,
    aes(full_species, mu, fill = species),
    slab_alpha = 0.5,
    .width = c(0.001, 0.95),
    point_interval = "mean_hdi"
  ) +
  geom_point(
    data = observed_species_mu,
    aes(full_species, mu), pch = 21, fill = "white", size = 2
  ) +
  scale_fill_manual(values = species_palette) +
  scale_x_discrete("", guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous("(fecundity|no retention) \n (nb eggs parasitized/2 days)")+
  annotate("text", x=1, y=35, label= "a") +
  annotate("text", x=2, y=60, label= "b") +
  annotate("text", x=3, y=40, label= "ac") +
  annotate("text", x=4, y=40, label= "ac") +
  annotate("text", x=5, y=50, label= "bc") +
  coord_cartesian(xlim = c(1, 5))

plot_fecundity

plot_zi <- ggplot() +
  stat_pointinterval(
    data = model1_data,
    aes(full_species, zi, group = line),
    col = "grey70",
    x = position_lines,
    .width = 0.95
  ) +
  stat_eye(
    data = fixed_data,
    aes(full_species, zi, fill = species),
    slab_alpha = 0.5,
    .width = c(0.001, 0.95),
    point_interval = "mean_hdi"
  ) +
  geom_point(
    data = observed_species_zi,
    aes(full_species, zi), pch = 21, fill = "white", size = 2
  ) +
  scale_fill_manual(values = species_palette) +
  scale_x_discrete("", guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous("retention probability") +
  annotate("text", x = 1, y = 0.65, label = "a") +
  annotate("text", x = 2, y = 0.75, label = "b") +
  annotate("text", x = 3, y = 0.65, label = "ab") +
  annotate("text", x = 4, y = 0.75, label = "ab") +
  annotate("text", x = 5, y = 0.85, label = "b") +
  coord_cartesian(xlim = c(1, 5))

plot_zi

# fuse all plots in one
fig2 <- (plot_dispersal / plot_development / plot_fecundity / plot_zi) &
  theme_bw() &
  theme(legend.position = "none", axis.text.x = element_markdown()) &
  plot_annotation(tag_levels = "A") 


# save generated figures in 'figures' folder, at the right dimensions
ggsave(here("figures", "fig2.pdf"), width = 163, height = 230, dpi = 1200, units = "mm")
ggsave(here("figures", "fig2.png"), width = 163, height = 230, dpi = 1200, units = "mm")
```

### Figure 3

WE want to see the correlation between fecundity and development time (Figure 3)

```{r figure POL}
# covariance inter and intra species; only on the mu component of fecundity = fecundity of individuals without retention

# first create a table for species highest density interval (fixed effect only)
figure_POL_mu <- fixed_data |>
  group_by(species) |>
  mean_hdi(mu) |> # 95 highest density interval (.width = 0.95)
  rename(mu.lower = .lower) |>
  rename(mu.upper = .upper)

figure_POL_development <- fixed_data |>
  group_by(species) |>
  mean_hdi(development_time) |> 
  rename(development.lower = .lower) |>
  rename(development.upper = .upper)

figure_POL_species <- figure_POL_development |>
  left_join(figure_POL_mu) |>
  mutate(full_species = paste0("*T. ", species, "*"))


# then a table for lines highest density intervals
figure_POL_mu <- model1_data |>
  group_by(species, line) |>
  mean_hdi(mu) |>
  rename(mu.lower = .lower) |>
  rename(mu.upper = .upper)

figure_POL_development <- model1_data |>
  group_by(species, line) |>
  mean_hdi(development_time) |>
  rename(development.lower = .lower) |>
  rename(development.upper = .upper)

figure_POL_line <- figure_POL_development |>
  left_join(figure_POL_mu)

# the plot
ggplot(data = figure_POL_line) +
  geom_linerange(
    aes(
      y = development_time, x = mu,
      ymin = development.lower, ymax = development.upper
    ),
    size = 0.5, alpha = 0.5, colour = "grey70"
  ) +
  geom_linerange(
    aes(
      y = development_time, x = mu,
      xmin = mu.lower, xmax = mu.upper
    ),
    size = 0.5, alpha = 0.5, colour = "grey70"
  ) +
  geom_point(aes(y = development_time, x = mu), size = 0.5) +
  geom_point(aes(y = development_time, x = mu), size = 0.5, colour = "grey70") +
  geom_linerange(
    data = figure_POL_species,
    aes(
      y = development_time, x = mu,
      ymin = development.lower, ymax = development.upper,
      col = full_species
    ), size = 1.1
  ) +
  geom_linerange(
    data = figure_POL_species,
    aes(
      y = development_time, x = mu,
      xmin = mu.lower, xmax = mu.upper,
      col = full_species
    ), size = 1.1
  ) +
  geom_point(data = figure_POL_species, mapping = aes(
    y = development_time, x = mu, fill = full_species,
    pch = full_species
  ), size = 4, colour = "black") +
  xlab("(Fecundity|no retention) (nb eggs parasitized/2 days)") +
  ylab("Development time (days)") +
  labs(fill = "species", col = "species", shape = "species") +
  scale_color_manual(values = species_palette) +
  scale_fill_manual(values = species_palette) +
  scale_shape_manual(values = 21:25) +
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.text = element_markdown()
  )

ggsave(here("figures", "fig3.pdf"), width = 152, height = 108, dpi = 1200, units = "mm")
ggsave(here("figures", "fig3.png"), width = 152, height = 108, dpi = 1200, units = "mm")
```

### Table 2: info for correlations tables

Now we look at line level correlations between our four traits, with and without species effect (Table 2)

```{r correlations}

# create summaries for the two models, with HDI credible intervals rather than the default quantile intervals of summary()

summary_model1 <- model1 |>
  as_draws_df() |>
  select(starts_with(c("b_", "sd_", "cor_", "sigma"))) |> # MAXIME: on peut se contenter de "cor" ici probablement (et en dessous) je supprime pas de suite au cas ou on a une idee qui utilise les autres coefs
  pivot_longer(everything()) |>
  group_by(name) |>
  mean_hdi()

summary_model1 |>
  print(n = Inf)

summary_model2 <- model2 |>
  as_draws_df() |>
  select(starts_with(c("b_", "sd_", "cor_", "sigma"))) |>
  pivot_longer(everything()) |>
  group_by(name) |>
  mean_hdi()

summary_model2 |>
  print(n = Inf)

# model 1: correlations partitioned into among-species and within species components,
# so the random effect correlation table only reflect within species correlation

table_cor_LINE <- summary_model1 |>
  mutate(value = paste(round(value, 2), " [", round(.lower, 2), "; ", round(.upper, 2), "]", sep = "")) |>
  filter(str_detect(name, "cor_line") == TRUE) |>
  mutate(
    response1 = case_when(
      str_detect(name, "cor_line__lowbound") ~ "lowbound",
      str_detect(name, "cor_line__zi") ~ "retention_probability",
      str_detect(name, "cor_line__nbegg") ~ "fecundity_noretention",
      str_detect(name, "cor_line__dispersal") ~ "dispersal_rate"
    ),
    response2 = case_when(
      str_detect(name, "lowbound_Intercept$") ~ "lowbound",
      str_detect(name, "zi_nbegg_Intercept$") ~ "retention_probability",
      str_detect(name, "nbegg_Intercept$") ~ "fecundity_noretention",
      str_detect(name, "dispersal_Intercept$") ~ "dispersal_rate"
    ),
  ) |>
  select(response1, response2, value)

## first model table
table_cor_LINE



# same but not partitioned into among-species and within species components = model 2

table_cor_fullLINE <- summary_model2 |>
  mutate(value = paste(round(value, 2), " [", round(.lower, 2), "; ", round(.upper, 2), "]", sep = "")) |>
  filter(str_detect(name, "cor_line") == TRUE) |>
  mutate(
    response1 = case_when(
      str_detect(name, "cor_line__lowbound") ~ "development_time",
      str_detect(name, "cor_line__zi") ~ "retention_probability",
      str_detect(name, "cor_line__nbegg") ~ "fecundity_noretention",
      str_detect(name, "cor_line__dispersal") ~ "dispersal_rate"
    ),
    response2 = case_when(
      str_detect(name, "lowbound_Intercept$") ~ "development_time",
      str_detect(name, "zi_nbegg_Intercept$") ~ "retention_probability",
      str_detect(name, "nbegg_Intercept$") ~ "fecundity_noretention",
      str_detect(name, "dispersal_Intercept$") ~ "dispersal_rate"
    ),
  ) |>
  select(response1, response2, value)


## second model table
table_cor_fullLINE
```

### BONUS JUST TO CHECK (you can delete later maybe if we don't use)

A version of figure 3, but for ZI and mu instead

```{r figure POL 2}
# covariance inter and intra species; only on the mu component of fecundity = fecundity of individuals without retention

# first create a table for species highest density interval (fixed effect only)
figure_POL_zi <- fixed_data |>
  group_by(species) |>
  mean_hdi(zi) |> # 95 highest density interval (.width = 0.95)
  rename(zi.lower = .lower) |>
  rename(zi.upper = .upper)

figure_POL_mu <- fixed_data |>
  group_by(species) |>
  mean_hdi(mu) |> # 95 highest density interval (.width = 0.95)
  rename(mu.lower = .lower) |>
  rename(mu.upper = .upper)

figure_POL_species_bonus <- figure_POL_zi |>
  left_join(figure_POL_mu) |>
  mutate(full_species = paste0("*T. ", species, "*"))


# then a table for lines highest density intervals
figure_POL_mu <- model1_data |>
  group_by(species, line) |>
  mean_hdi(mu) |>
  rename(mu.lower = .lower) |>
  rename(mu.upper = .upper)

figure_POL_zi <- model1_data |>
  group_by(species, line) |>
  mean_hdi(zi) |>
  rename(zi.lower = .lower) |>
  rename(zi.upper = .upper)

figure_POL_line_bonus <- figure_POL_zi |>
  left_join(figure_POL_mu)

# the plot
ggplot(data = figure_POL_line_bonus) +
  geom_linerange(
    aes(
      y = zi, x = mu,
      ymin = zi.lower, ymax = zi.upper
    ),
    size = 0.5, alpha = 0.5, colour = "grey70"
  ) +
  geom_linerange(
    aes(
      y = zi, x = mu,
      xmin = mu.lower, xmax = mu.upper
    ),
    size = 0.5, alpha = 0.5, colour = "grey70"
  ) +
  geom_point(aes(y = zi, x = mu), size = 0.5) +
  geom_point(aes(y = zi, x = mu), size = 0.5, colour = "grey70") +
  geom_linerange(
    data = figure_POL_species_bonus,
    aes(
      y = zi, x = mu,
      ymin = zi.lower, ymax = zi.upper,
      col = full_species
    ), size = 1.1
  ) +
  geom_linerange(
    data = figure_POL_species_bonus,
    aes(
      y = zi, x = mu,
      xmin = mu.lower, xmax = mu.upper,
      col = full_species
    ), size = 1.1
  ) +
  geom_point(data = figure_POL_species_bonus, mapping = aes(
    y = zi, x = mu, fill = full_species,
    pch = full_species
  ), size = 4, colour = "black") +
  xlab("(Fecundity|no retention) (nb eggs parasitized/2 days)") +
  ylab("retention rate") +
  labs(fill = "species", col = "species", shape = "species") +
  scale_color_manual(values = species_palette) +
  scale_fill_manual(values = species_palette) +
  scale_shape_manual(values = 21:25) +
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.text = element_markdown()
  )
```

## BONUS 2 suitable for supplementary?

```{r plot trait values line}
## plot posterior traits values per species (fixed effects only), with posterior lines values as boxplot
model1_data$full_species <- paste0("*T. ", model1_data$species, "*")

## observed means for plotting (approximated for development and fecundity variables)
observed_species_disp <- data_Trichogramma |>
  filter(valid_dispersal == 1) |>
  group_by(species,line) |>
  summarise(disp = mean(dispersal)) |>
  mutate(full_species = paste0("*T. ", species, "*"))

observed_species_dev <- data_Trichogramma |>
  filter(valid_development == 1) |>
  group_by(species,line) |>
  summarise(time = mean((upbound + lowbound) / 2)) |>
  mutate(full_species = paste0("*T. ", species, "*"))

observed_species_zi <- data_Trichogramma |>
  filter(valid_fecundity == 1) |>
  group_by(species,line) |>
  summarise(zi = mean(nb_egg == 0)) |>
  mutate(full_species = paste0("*T. ", species, "*"))

observed_species_mu <- data_Trichogramma |>
  filter(valid_development == 1 & nb_egg > 0) |>
  group_by(species,line) |>
  summarise(mu = mean(nb_egg)) |>
  mutate(full_species = paste0("*T. ", species, "*"))



plot_dispersal <- ggplot() +
  stat_pointinterval(
    data = model1_data,
    aes(line, dispersal_rate, col = full_species),
    .width = 0.95
  ) +
  geom_point(
    data = observed_species_disp,
    aes(line, disp), pch = 21, fill = "white", size = 2
  ) +
  scale_color_manual(values = species_palette) +
  scale_x_discrete("", guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous("dispersal rate")+ 
  facet_grid(cols=vars(full_species),scale = "free_x")

plot_dispersal


plot_development <- ggplot() +
  stat_pointinterval(
    data = model1_data,
    aes(line, development_time, col=full_species),
    .width = 0.95
  ) +
  geom_point(
    data = observed_species_dev,
    aes(line, time), pch = 21, fill = "white", size = 2
  ) +
  scale_color_manual(values = species_palette) +
  scale_x_discrete("", guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous("development time (days)")+ 
  facet_grid(cols=vars(full_species),scale = "free_x")

plot_development

plot_fecundity <- ggplot() +
  stat_pointinterval(
    data = model1_data,
    aes(line, mu, col=full_species),
    .width = 0.95
  ) +
  geom_point(
    data = observed_species_mu,
    aes(line, mu), pch = 21, fill = "white", size = 2
  ) +
  scale_color_manual(values = species_palette) +
  scale_x_discrete("", guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous("(fecundity|no retention) \n (nb eggs parasitized/2 days)")+ 
  facet_grid(cols=vars(full_species),scale = "free_x")

plot_fecundity

plot_zi <- ggplot() +
  stat_pointinterval(
    data = model1_data,
    aes(line, zi, col=full_species),
    .width = 0.95
  ) +
  geom_point(
    data = observed_species_zi,
    aes(line, zi), pch = 21, fill = "white", size = 2
  ) +
  scale_color_manual(values = species_palette) +
  scale_x_discrete("", guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous("retention rate")+ 
  facet_grid(cols=vars(full_species),scale = "free_x")

plot_zi

(plot_dispersal / plot_development / plot_fecundity / plot_zi) &
  theme_bw() &
  theme(legend.position = "none", strip.text.x = element_textbox()) &
  plot_annotation(tag_levels = "A")
```
